---
title: "Formal Time Series Models"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

## AR(1) process
The R function `arima.sim` lets us simulate ARIMA processes.

```{r, flcolours=TRUE}
x <- arima.sim(list(ar=c(0.75)), n=100)
plot(x, xlab="Time t", ylab=expression(X[t]), col=2)
title("MA(1) time series")
```
The autocorrelation function of an MA(1) drops to 0 after lag 1.
```{r, flcolours=TRUE}
acf(x)
```
The *partial* autocorrelation function of an AR(1) process decays drops to zero after lag 1.
```{r, flcolours=TRUE}
pacf(x)
```
## MA(1) process

```{r, flcolours=TRUE}
x <- arima.sim(list(ma=1), n=100)
plot(x, xlab="Time t", ylab=expression(X[t]), col=2)
title("MA(1) time series")
```
The autocorrelation function of an MA(1) drops to zero after lag 1.
```{r, flcolours=TRUE}
acf(x)
```

The partial autocorrelation function of an MA(1) oscillates and slowly decays to zero.
```{r, flcolours=TRUE}
pacf(x)
```


## Code used for the simulation


```{r, flcolours=TRUE}
t <- 1:100/10
x <- arima.sim(list(ma=1), n=100) + t
plot(x, xlab="Time t", ylab=expression(X[t]), col=2)
abline(0,0.1, lty=3, col=3)
title("MA(1) time series with trend")
```


```{r, flcolours=TRUE}
x <- arima.sim(list(ma=1), n=100) + t +3*sin(t/3*2*pi) + 2*cos(t/3*2*pi)
plot(x, xlab="Time t", ylab=expression(X[t]), col=2)
title("MA(1) time series with trend and seasonal component")
s <- seq(0, 101, len=1e3)
lines(s, s/10+3*sin(s/30*2*pi) + 2*cos(s/30*2*pi), lty=3, col=3)
```



