---
title: "Autocorrelation"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

## Purely random data 

We first generate the data as Gaussian white noise.
```{r}
set.seed(1)
x <- rnorm(100, mean=0, sd=1)
```
We can then plot the data ..
```{r, flcolours=TRUE}
plot(x, type="l", col=2, xlab="", ylab="", main="Time Series")
```
... and the autocorrelation function.
```{r, flcolours=TRUE}
acf(x, lag=25, main="Autocorrelation function")
```

## Data with a trend

We first generate the data and plot it.
```{r, flcolours=TRUE}
set.seed(2)
time <- 1:100
x <- -0.1*time + rnorm(100, mean=0, sd=1)
plot(x, type="l", col=2, xlab="", ylab="", main="Time Series")
```
We can then generate the autocorrelation function.
```{r, flcolours=TRUE}
acf(x, lag=25, main="Autocorrelation function")
```
The linear trend leads to a very slowly decaying autocorrelation function.

## Data with a seasonal effect

We first generate data with a seasonal, in our case cosine, component and plot it
```{r, flolours=TRUE}
set.seed(3)
time <- 1:100
seasonal <- 2*cos(time/(0.5*pi)/12*10)
x <- seasonal + rnorm(100, mean=0, sd=1)
plot(x, type="l", col=2, xlab="", ylab="", main="Time Series")
```
The autocorrelation function now shows a peak at the period (12) of the seasonal effect.
```{r, flcolours=TRUE}
acf(x, lag=25, main="Autocorrelation function")
```

## Data with a trend and a seasonal effect

We finally add both a trend and seasonal component to the simulated time series.
```{r, flcolours=TRUE}
set.seed(4)
time <- 1:100
seasonal <- 2*cos(time/(0.5*pi)/12*10)
x <- seasonal + rnorm(100, mean=0, sd=1) + 0.1*time
plot(x, type="l", col=2, xlab="", ylab="", main="Time Series")
```
The autocorrelation function now shows both a slow decay and an additional peak at 12 corresponding to the period of the seasonal effect.
```{r, flcolours=TRUE}
acf(x, lag=25, main="Autocorrelation function")
```