---
title: "Temperatures in Eskdalemuir"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

## Temperatures in Eskdalemuir

We start by loading the Eskdalemuir data, which we have already used this weeek, and plot the time series again.
```{r, flcolours=TRUE}
load("../../data/week2/esktem.RData")
plot(esktem, main="Temperatures in Eskdalemuir", col=2)
```

Next, we have to construct the covariate matrix containing the linear trend and the harmonic terms for the seasonal effect.
```{r}
t <- seq(1914, length=length(esktem), by=1/12)
xreg <- cbind(t, sin(2*pi*t), cos(2*pi*t))
```

We can now use the function `auto.arima` to search for the best model.
```{r}
library(forecast)
model <- auto.arima(esktem, seasonal=FALSE, xreg=xreg)
summary(model)
```
Next we plot the linear trend and overlay the partial residuals (i.e. we remove the estimated seasonal effect from the temperatures to obtain what is effectively a temperature anomaly).

```{r, flcolours=TRUE}
beta <- coef(model)[5:8]
plot(t, esktem-beta[3]*sin(2*pi*t) - beta[4]*cos(2*pi*t), col=3, xlab="Year", ylab="Long-term trend")
title("Long-term trend")
abline(beta[1], beta[2], lwd=2, col=2)
```

Finally we plot the seasonal effect and overlay the partial residuals (this time ew remove the long-term trend).

```{r, flcolours=TRUE}
plot(t%%1, esktem-beta[1]-beta[2]*t, xaxt="n", xlab="Month", ylab="Seasonal effect", col=3)
title("Seasonal effect")
axis(side=1, at=0:11/12,labels=c("Jan","Feb","Mar","Apr","May", "Jun", "Jul", "Aug","Sep","Oct", "Nov","Dec"))
tt <- seq(0,1,len=1e4)
lines(tt, beta[3]*sin(2*pi*tt) + beta[4]*cos(2*pi*tt), lwd=2, col=2)
```