---
title: "Forecasting"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

We start by loading the NO<sub>2</sub> data.
```{r}
load("../../data/week2/glasgowno2.RData")
```

We start by plotting the data.
```{r, flcolours=TRUE}
plot(no2, axes=FALSE, ylab=expression(NO[2]),xlim=c(1,9), col=2)          
axis(2)
axis(1, labels=c("31/01", "01/02", "02/02", "03/02", "04/02", "05/02", "06/02", "07/02", "08/02"), at=1:9)
```

Next, we fit a time series model using a trend and diurnal ("seasonal") component. The diurnal pattern looks more complex than a straight sine curve, so we will also include higher frequency terms.
```{r}
t <- 1:length(no2)
xreg <- cbind(t, sin(2*pi*t/24), cos(2*pi*t/24),  sin(2*pi*t/12), cos(2*pi*t/12), 
              sin(2*pi*t/6), cos(2*pi*t/6))
model <- arima(no2, order=c(2,0,0), xreg=xreg)
summary(model)
```
We can now add the 
```{r, flcolours=TRUE}
plot(no2, axes=FALSE, ylab=expression(NO[2]),xlim=c(1,9), col=2)          
axis(2)
axis(1, labels=c("31/01", "01/02", "02/02", "03/02", "04/02", "05/02", "06/02", "07/02", "08/02"), at=1:9)
lines(1+t/24, cbind(1,xreg)%*%coef(model)[3:10], col=3)
```


To forecast for the next 52 hours we have to create the `xreg` matrix corresponding to $t=142,\ldots,192$.
```{r}
newt <- length(no2)+1:52
newxreg <- cbind(newt, sin(2*pi*newt/24), cos(2*pi*newt/24),  sin(2*pi*newt/12),  
                 cos(2*pi*newt/12), sin(2*pi*newt/6), cos(2*pi*newt/6))
forecast <- predict(model, n.ahead=52, newxreg=newxreg)
```

We can now add the forecast to the plot.
```{r, flcolours=TRUE}
plot(no2, axes=FALSE, ylab=expression(NO[2]),xlim=c(1,9), col=2)          
axis(2)
axis(1, labels=c("31/01", "01/02", "02/02", "03/02", "04/02", "05/02", "06/02", "07/02", "08/02"), at=1:9)
lines(1+t/24, cbind(1,xreg)%*%coef(model)[3:10], col=3)
lines(1+newt/24, forecast$pred, col=4)
abline(v=1+140/24, lty=2)
```
