---
title: "Introductory Example: Atmospheric Carbon Dioxide at Manua Loa"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

## Data

We start by loading the Mauna Loa data into R.
```{r}
loadata <- read.csv("../../data/week2/maunaloa.csv")
head(loadata)
```

At this stage we are interested in the column `average`, which we will now convert in to a `ts` object, the data structure for time series in R. 

```{r}
loadata <- ts(loadata$average, start=c(1958,3), frequency=12)
loadata
```
We will now save this as an `.RData` files, so we can go straight to the `ts` object when we analyse the time series later on.

```{r}
save(loadata, file="../../data/week2/maunaloa.RData")
```

Now we can plot the data 
```{r}
plot(loadata)
```

On FutueLearn we have use a different colour (`col=2`), points rather than lines (`type="l"`), a different plotting symbol (`pch=15`) and a smaller plotting symbol (`cex=0.5`).
```{r}
plot(loadata, col=2, type="p", pch=16, cex=0.5,
     xlab="Year", ylab="Carbon dioxide level (ppmnv)")
title(expression("Mauna Loa CO"[2]*" levels"))
```

Too better see the periodic nature of the time series we zoom into the period from 2010 to 2018. We do thsi by specifying `xlim=c(2010,2018)`.
```{r}
plot(loadata, col=2, type="p", pch=16, cex=0.5, 
     xlim=c(2010, 2018),
     xlab="Year", ylab="Carbon dioxide level (ppmnv)")
title(expression("Mauna Loa CO"[2]*" levels"))
```

Alternatively we can subset the time series object using teh function `subset` from the pakcage `forecast`.
```{r}
library(forecast)
loashort <- subset(loadata, start=623, end=718)
plot(loashort, col=2, type="p", pch=16, cex=0.5,
     xlab="Year", ylab="Carbon dioxide level (ppmnv)")
title(expression("Mauna Loa CO"[2]*" levels"))
```