---
title: "Introductory Example: Wind speed"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

## Data

We start by loading the Wind speed data into R and convert the column `Newdate` into R's `Date` format.
```{r}
site1 <- read.csv("../../data/week2/windspeed.csv")
site1$Newdate <- as.Date(site1$Newdate)
```

We can now plot the data (we'll use the formula notation `y~x` to specify what which variables to use for the x and y axes).
```{r, flcolours=TRUE}
plot(WindSpeed~Newdate, data=site1, ylab="Wind Speed", xlab="Year", 
     pch=19, cex=0.2, col=2)
title("Wind Speed")
```

To reduce the clutter we can calculate the average wind speed for every day. This can be done very efficiently using the function `aggregate`.

```{r}
# Calculate dailCon vert the date to a date#Wind speed
agg_mean <- aggregate(site1$WindSpeed, by=list(Date=site1$Newdate), mean)
head(agg_mean)
```

We can now plot the daily average wind speed as a line
```{r, flcolours=TRUE}
plot(x~Date, data=agg_mean, ylab="Wind Speed", xlab="Year", 
     type="l", col=2)
title("Wind Speed")
```

