---
title: "Wind speeds in Claremorris"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../theme/theme-mooc.R")
```

d 

We start with loading the data.
```{r, flcolours=TRUE}
load("../../data/week3/wind.RData")
plot(wind, xlab="Year", ylab="Wind speed in knots")
```

We start by looking for changes in the mean and the variance.
```{r, flcolours=TRUE}
library(changepoint)
par(mfrow=c(1,2))
plot(cpt.meanvar(wind, method="PELT", penalty="Hannan-Quinn"))
title("Hannan-Quinn penalty")
plot(cpt.meanvar(wind, method="PELT", penalty="BIC"))
title("BIC penalty")
```

We obtain a lot of changepoint, to a large extent due to the high autocorrelation in the data. We can try to reduce this autocorrelation by looking at the differences between observations, rather than the original data.

A plot of the differences is shown in the plot below.

```{r, flcolours=TRUE}
plot(diff(wind), xlab="Year", ylab="Wind speed differences in knots")
```

The changepoint model now appears to give more a more realistic picture of the changepoints.

```{r, flcolours=TRUE}
library(changepoint)
par(mfrow=c(1,2))
plot(cpt.var(diff(wind), method="PELT", penalty="Hannan-Quinn"))
title("Hannan-Quinn penalty (differences)")
plot(cpt.var(diff(wind), method="PELT", penalty="BIC"))
title("BIC penalty (differences)")
```

